02-preliminary-analysis

### 2024-05-28 data/01-census-and-traveltimes.qmd
Added more variables to "HAM_census_21.rda" using 01-census-and-traveltimes.qmd. Played around in 03-clustering-and-profile.qmd

```{r, eval=FALSE}
HAM_census_21 <- get_census(dataset='CA21', 
                            regions=list(CSD="3525005"),
                            vectors=c("v_CA21_11", #total - aged 0-14
                                      "v_CA21_12", #males - aged 0-14
                                      "v_CA21_13", #females - aged 0-14
                                      "v_CA21_68", #total - aged 15-64
                                      "v_CA21_69", #males - aged 15-64
                                      "v_CA21_70", #females - aged 15-64
                                      "v_CA21_251", #total - aged 65 and over
                                      "v_CA21_252", #males - aged 65 and over
                                      "v_CA21_253", #females - aged 65 and over
                                      "v_CA21_1085", #total - prevalence of LICO-AT (%)
                                      "v_CA21_1086", #male - prevalence of LICO-AT (%)
                                      "v_CA21_1087",#female - prevalence of LICO-AT (%)
                                      "v_CA21_7647", #total - walked to work
                                      "v_CA21_7635", #total - drove or was driven to work
                                      "v_CA21_7644", #total - used transit to get to work
                                      "v_CA21_7650", #total - biked to work
                                      "v_CA21_7653" #total - used some other mode to get to work
                                      ), 
                                      level='DA', use_cache = FALSE, geo_format = 'sf', quiet = TRUE)

>>>>

```{r, eval=FALSE}
HAM_census_21 <- get_census(dataset='CA21', 
                            regions=list(CSD="3525005"),
                            vectors=c("v_CA21_11", #total - aged 0-14
                                      "v_CA21_68", #total - aged 15-64
                                      "v_CA21_251", #total - aged 65 and over

                                       "v_CA21_498", #Average number of children in census families with children
                                      "v_CA21_500", # No. of couple families
                                      "v_CA21_507", # No. of (non-couple) one-parent families
                                      "v_CA21_1085", #Prevalence of LICO-AT (%)

                                      "v_CA21_1103", #No. of pop in priv households in the Adjusted after-tax family income decile group in the BOTTOM HALF of distribution
                                      "v_CA21_1121", #No. of pop in priv households in the Adjusted after-tax family income decile group in the TOP HALF of distribution

                                      "v_CA21_1142", #Gini index on adjusted household after-tax income
                                      "v_CA21_907", #Median after-tax income of household in 2020 ($)

                                      "v_CA21_945", # no. of households making under $5,000 (AT income)
                                      "v_CA21_946", # no. of households making $5000-$9999 (AT income)
                                      "v_CA21_947", # no. of households making $10000-$14999 (AT income)
                                      "v_CA21_948", # no. of households making $15000-$19999 (AT income)
                                      "v_CA21_949", # no. of households making $20000-$24999 (AT income)
                                      "v_CA21_950", # no. of households making $25000-$29999 (AT income)
                                      "v_CA21_951", # no. of households making $30000-$34999 (AT income)
                                      "v_CA21_952", # no. of households making $35000-$39999 (AT income)
                                      "v_CA21_953", # no. of households making $40000-$44999 (AT income)
                                      "v_CA21_954", # no. of households making $45000-$49999 (AT income)
                                      "v_CA21_955", # no. of households making $50000-$59999 (AT income)
                                      "v_CA21_956", # no. of households making $60000-$69999 (AT income)
                                      "v_CA21_957", # no. of households making $70000-$79999 (AT income)
                                      "v_CA21_958", # no. of households making $80000-$89999 (AT income)
                                      "v_CA21_959", # no. of households making $90000-$99999 (AT income)
                                      "v_CA21_960", # no. of households making over $100000 (AT income)

                                      "v_CA21_6498",# No. of pop. who are employed (in the labour force)
                                      "v_CA21_6501",#  No. of pop. who are not employed (in the labour force)
                                      "v_CA21_6504",#  No. of pop. who are not in the labour force

                                      "v_CA21_1144",# TOTAL - knowledge of official languages for the total population excluding institutional residents
                                       "v_CA21_1156",# No. of pop population with no knowledge of official languages (either English or French), NOTE: _1144 - 1156 = the No. of pop who have knowledge of at least 1 off. lang.

                                      "v_CA21_4875",# No. of pop who are visible minority
                                      "v_CA21_4914",# No. of pop who are are NOT visible minority

                                      "v_CA21_5868",#  No. of pop w/ No certificate, diploma or degree for the population aged 25 to 64 years in private households
                                      "v_CA21_5871",#  No. of pop (age 25 to 64 year) w/ Highest degree is a High (secondary) school diploma or equivalency certificate
                                      "v_CA21_5874",# No. of pop (age 25 to 64 year) w/ Highest degree is Postsecondary certificate, diploma or degree (bachelor or higher)

                                      "v_CA21_7647", # % total - walked to work
                                      "v_CA21_7635", # % total - drove or was driven to work
                                      "v_CA21_7644", # % total - used transit to get to work
                                      "v_CA21_7650", # % total - biked to work
                                      "v_CA21_7653" # % total - used some other mode to get to work
                                      ), 
                                      level='DA', use_cache = FALSE, geo_format = 'sf', quiet = TRUE)

### 2024-05-28 data/02-access-and-prelim-plots.qmd
Added more variables to "HAM_census_21.rda" using 01-census-and-traveltimes.qmd. Played around in 03-clustering-and-profile.qmd

```{r library-setup}
library(biscale)
library(cowplot)
library(sf)
library(here)
library(dplyr)
library(tmap)
library(mapview)
library(tmaptools)
library(pander)
library(flextable)
library(ftExtra)
library(scales)
```

@@

```{r}
HAM_census_21 <- HAM_census_21 |>
  rename("LICO_AT_TOTAL" = `v_CA21_1085: Prevalence of low income based on the Low-income cut-offs, after tax (LICO-AT) (%)`) |> 
  mutate(LICO_AT_TOTAL = ifelse(is.na(LICO_AT_TOTAL), 0, LICO_AT_TOTAL)) |>
  rowwise() |> mutate(
         perc_walked = (`v_CA21_7647: Walked`)/sum(`v_CA21_7647: Walked`,`v_CA21_7635: Car, truck or van`, `v_CA21_7644: Public transit`,`v_CA21_7650: Bicycle`, `v_CA21_7653: Other method`, na.rm=TRUE),
         perc_car = (`v_CA21_7635: Car, truck or van`)/sum(`v_CA21_7647: Walked`,`v_CA21_7635: Car, truck or van`, `v_CA21_7644: Public transit`,`v_CA21_7650: Bicycle`, `v_CA21_7653: Other method`, na.rm=TRUE),
perc_transit = (`v_CA21_7644: Public transit`)/sum(`v_CA21_7647: Walked`,`v_CA21_7635: Car, truck or van`, `v_CA21_7644: Public transit`,`v_CA21_7650: Bicycle`, `v_CA21_7653: Other method`, na.rm=TRUE),
 perc_cycle_other = (`v_CA21_7650: Bicycle`+ `v_CA21_7653: Other method`)/sum(`v_CA21_7647: Walked`,`v_CA21_7635: Car, truck or van`, `v_CA21_7644: Public transit`,`v_CA21_7650: Bicycle`, `v_CA21_7653: Other method`, na.rm=TRUE)) |>
  select(c("GeoUID", "Population", "Households", "Dwellings", "LICO_AT_TOTAL", "perc_walked", "perc_car", "perc_transit", "perc_cycle_other"))

@@

Aggregating time!
```{r}
#ID = care destination ID (in care_dest), GeoUID is id of the DA, from_id is parcel ID
save(access_15, file = "access_15.rda", compress = "xz")
```

@@

```{r}
#by DA and mean:
access_15_DA <- access_15 |>
  left_join(HAM_census_21 |> st_drop_geometry() |> select(c("GeoUID", "Population", "Dwellings", "LICO_AT_TOTAL", "COMMUNITY_")), 
            by=c("GeoUID")) |> 
  group_by(GeoUID) |>
  summarise(copp_iscDA_med = median(copp_isc), #the median number of opportunities (all) that can be reached from each parcel in a DA
            copp_iscDA_sd = sd(copp_isc), #the SD
    copp_iscDA_w_med = median(copp_isc_w), #the median number of opportunities (all) that can be reached from each parcel in a DA
            copp_iscDA_w_sd = sd(copp_isc_w), #the SD
            V_iscDA = sum(V_isc), #the number of total spatially available opportunities (all) that can be reached from each parcel in a DA
            vv_iscDA = sum(V_isc)/first(Population), #the per-capita number of total spatially available opportunities (all) that can be reached from each parcel in a DA
            Population = first(Population),
            Dwellings = first(Dwellings),
            LICO_AT_TOTAL = first(LICO_AT_TOTAL),
            COMMUNITY_ = first(COMMUNITY_)) |>
  tidyr::drop_na(Population)

@@

```{r}
#calculating entropy, here I do it by the 14 specific categories. This is a way to consider their diversity within and between categories. Because they have been weighted before calc. the cumulative opp counts, we can proceed without anymore adjustments.
step1 <- access_15_byspecificcategory_DA  |> group_by(GeoUID) |> 
  summarise(OPPi = sum(copp_isc_med, na.rm=T) + sum(copp_isc_sd, na.rm=T),
            Vi = sum(V_isc))

step2 <- access_15_byspecificcategory_DA |> left_join(step1, by="GeoUID")

step2 <- step2 |> mutate(top1_copp = (copp_isc_med+copp_isc_sd) / OPPi * log((copp_isc_med+copp_isc_sd) / OPPi),
             top1_copp = ifelse(is.nan(top1_copp), 0, top1_copp),
             top1_V= (V_isc) / Vi * log((V_isc) / Vi),
             top1_V = ifelse(is.nan(top1_V), 0, top1_V),)  |> 
  group_by(GeoUID) |>
  summarise(entropy_copp_isc_medsd = sum(top1_copp) / -log(14),
            entropy_V_isc = sum(top1_V) / -log(14))

@@

```{r}
#the number of opportunities from each categories
all_cats <- access_15_bycategory_DA |> 
  mutate(copp_all = sum(copp_ic_med, na.rm=T) + sum(copp_ic_sd, na.rm=T)) |>
  group_by(GeoUID) |>
         summarise(summed_copp_all = sum(copp_all),
                   summed_Vic = sum(V_ic, na.rm=T),
                   Population = mean(Population),
                   Care_Access_Status1 = case_when(summed_copp_all < 4 ~ "[0-04 des.) Not caring",
                                 summed_copp_all >= 4 & summed_copp_all < 25 ~ "[04-25 des.) Caring", 
                                 summed_copp_all >= 25 ~ "[25+ des.) Very caring"),
                   Care_Access_Status2 = case_when(all(copp_all >= 2 & copp_all < 4) ~ "- complete (2 to 4 options per cat.)", 
                                                   all(copp_all >= 4) ~ "- very complete (4 or more options per cat.)"),
                   entropy_copp_isc_medsd = first(entropy_copp_isc_medsd),
                   entropy_V_isc = first(entropy_V_isc)) |>
  mutate(Care_Access_Status2 = ifelse(is.na(Care_Access_Status2), "", Care_Access_Status2),
         Care_Access_Status = paste(Care_Access_Status1,Care_Access_Status2, sep=" "),
         vvi = summed_Vic/Population)

@@

```{r message=FALSE}
#tmap_mode("plot")
all_cats_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(all_cats) +
  tm_polygons("Care_Access_Status",
              border.alpha = 0.2,
              palette = c("darkred", "orangered", "orange", "yellow", "forestgreen", "lightgreen"),
              title = "Number of care destinations (all) spatially accessible within 15 min. walk",
              legend.is.portrait = TRUE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("right", "top"), size=1.0)+
  tm_layout(bg.color = "grey",
            panel.label.bg.color = "white",
            panel.label.color = "black",
            legend.outside = TRUE,
            legend.position = c("center", "bottom"),
            legend.outside.position = "bottom",
            legend.outside.size = 0.2,
            legend.width = 12
            # legend.title.size = 0.7,
            # legend.text.size = 0.6
            )

tmap_save(all_cats_copp_plot, file="all_cats_copp_plot.tiff")
```

```{r}
all_cats_entropy_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(all_cats) +
  tm_polygons("entropy_copp_isc_medsd",
              border.alpha = 0.2,
              palette = "Reds",
              title = "Diversity of (all) care destinations spatially accessible within 15 min. walk",
              legend.is.portrait = TRUE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("right", "top"), size=1.0)+
  tm_layout(bg.color = "grey",
            panel.label.bg.color = "white",
            panel.label.color = "black",
            legend.outside = TRUE,
            legend.position = c("center", "bottom"),
            legend.outside.position = "bottom",
            legend.outside.size = 0.2,
            legend.width = 12)

tmap_save(all_cats_entropy_copp_plot, file="all_cats_entropy_copp_plot.tiff")
```

```{r}
all_cats_vvi_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(all_cats) +
  tm_polygons("vvi",
              border.alpha = 0.2,
              palette = c("orangered", "forestgreen"),
              breaks = c(0, 0.003804, 0.022174),
              labels = c("Below Median", "Median and above"),
              title = "Per capita spatially available care destinations (all - weighted) within 15 min. walk",
              legend.is.portrait = TRUE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("right", "top"), size=1.0)+
  tm_layout(bg.color = "grey",
            panel.label.bg.color = "white",
            panel.label.color = "black",
            legend.outside = TRUE,
            legend.position = c("center", "bottom"),
            legend.outside.position = "bottom",
            legend.outside.size = 0.2,
            legend.width = 12
            )

tmap_save(all_cats_vvi_plot, file="all_cats_vvi_plot.tiff")
```

@@

```{r}
#adding entropy labels
all_cats <- all_cats |> 
  mutate(bivariate_entropy = 
           case_when( ((Care_Access_Status == "[04-25 des.) Caring - complete (2 to 4 options per cat.)"  |
                       Care_Access_Status =="[04-25 des.) Caring - very complete (4 or more options per cat.)" |
                         Care_Access_Status =="[25+ des.) Very caring - very complete (4 or more options per cat.)") &
                         entropy_copp_isc_medsd >= 0.7418) ~ "Caring, complete and amenity diverse",  #the median diversity,

                     ((Care_Access_Status == "[04-25 des.) Caring - complete (2 to 4 options per cat.)"  |
                       Care_Access_Status =="[04-25 des.) Caring - very complete (4 or more options per cat.)" |
                         Care_Access_Status =="[25+ des.) Very caring - very complete (4 or more options per cat.)") &
                        entropy_copp_isc_medsd < 0.7418) ~ "Caring, complete but not amenity diverse",

                     ((Care_Access_Status == "[0-04 des.) Not caring " |
                       Care_Access_Status == "[04-25 des.) Caring ")  &
                        entropy_copp_isc_medsd >= 0.7418) ~ "Not caring or not complete, but amenity diverse",

                     ((Care_Access_Status == "[0-04 des.) Not caring "  |
                       Care_Access_Status == "[04-25 des.) Caring ")  &
                        entropy_copp_isc_medsd < 0.7418) ~ "Not caring, complete, or amenity diverse"))

@@

```{r}
all_cats_bivariate_entropy_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(all_cats) +
  tm_polygons("bivariate_entropy",
              border.alpha = 0.2,
              palette = c("yellowgreen","yellow","blue", "#FFFFDF"),
              title = "Diversity of (all) care destinations spatially accessible within 15 min. walk",
              legend.is.portrait = TRUE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("right", "top"), size=1.0)+
  tm_layout(bg.color = "grey",
            panel.label.bg.color = "white",
            panel.label.color = "black",
            legend.outside = TRUE,
            legend.position = c("center", "bottom"),
            legend.outside.position = "bottom",
            legend.outside.size = 0.2,
            legend.width = 12)

tmap_save(all_cats_bivariate_entropy_copp_plot, file="all_cats_bivariate_entropy_copp_plot.tiff")
```

```{r}
care_categories_plot <- 
 tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(Community_Boundaries, bbox=Community_Boundaries) + tm_polygons(col = "white", border.col = "black", lwd=1)+
  tm_shape(care_dest) + tm_dots("Care_Category_Specific", palette = RColorBrewer::brewer.pal(n = 4, name = 'Dark2'), size=0.025, title = "")+
  tm_facets("Care_Category", nrow = 2, free.scales= TRUE)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 20)) +
  tm_compass(position = c("left", "top"), size=0.9) +
  tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 0.5, legend.bg.color = "white", legend.bg.alpha = 0.4,
            panel.labels = c("1,265 Dependent-centric destinations", 
                             "158 Errand-centric destinations", 
                             "381 Grocery-centric destinations", 
                             "421 Health-centric destinations"),
            panel.label.size = 0.8) 
care_categories_plot

tmap_save(care_categories_plot, filename ="care_categories_plot.tiff")
```
```{r}
parcels_DA_plot <- 
 tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(access_15_DA) +
  tm_polygons("parcel_count",
              border.alpha = 0.2,
              palette = "Reds",style = "quantile", n=4,
              title = "Count of residential parcels",
              legend.is.portrait = TRUE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("left", "top"), size=1.0)+
   tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 0.5,legend.title.size = 0.7,  legend.bg.color = "white", legend.bg.alpha = 0.4)

@@

pop_descriptive_plots <- tmap_arrange(parcels_DA_plot, poprate_DA_plot, LICOAT_DA_plot, nrow=3)
tmap_save(pop_descriptive_plots, file="pop_descriptive_plots.tiff", asp = 1/0.6)
```

>>>>

```{r library-setup}
library(biscale)
library(cowplot)
library(sf)
library(here)
library(dplyr)
library(tmap)
library(mapview)
library(tmaptools)
library(pander)
library(flextable)
library(ftExtra)
library(scales)
library(here)
```

@@

```{r}
HAM_census_21 <- HAM_census_21 |>
  rename("LICO_AT_TOTAL" = `v_CA21_1085: Prevalence of low income based on the Low-income cut-offs, after tax (LICO-AT) (%)`) |> 
  mutate(LICO_AT_TOTAL = ifelse(is.na(LICO_AT_TOTAL), 0, LICO_AT_TOTAL)) |>
  rowwise() |> mutate(
         perc_walked = (`v_CA21_7647: Walked`)/sum(`v_CA21_7647: Walked`,`v_CA21_7635: Car, truck or van`,`v_CA21_7644: Public transit`,`v_CA21_7650: Bicycle`, `v_CA21_7653: Other method`, na.rm=TRUE),
         perc_car = (`v_CA21_7635: Car, truck or van`)/sum(`v_CA21_7647: Walked`,`v_CA21_7635: Car, truck or van`, `v_CA21_7644: Public transit`,`v_CA21_7650: Bicycle`, `v_CA21_7653: Other method`, na.rm=TRUE),
perc_transit = (`v_CA21_7644: Public transit`)/sum(`v_CA21_7647: Walked`,`v_CA21_7635: Car, truck or van`, `v_CA21_7644: Public transit`,`v_CA21_7650: Bicycle`, `v_CA21_7653: Other method`, na.rm=TRUE),
 perc_cycle_other = (`v_CA21_7650: Bicycle`+ `v_CA21_7653: Other method`)/sum(`v_CA21_7647: Walked`,`v_CA21_7635: Car, truck or van`, `v_CA21_7644: Public transit`,`v_CA21_7650: Bicycle`, `v_CA21_7653: Other method`, na.rm=TRUE)) |>
  select(c("GeoUID", "Population", "Households", "Dwellings", "LICO_AT_TOTAL", "perc_walked", "perc_car", "perc_transit", "perc_cycle_other"))

@@

Aggregating time!
```{r}
#ID = care destination ID (in care_dest), GeoUID is id of the DA, from_id is parcel ID
save(access_15, file= file.path(here("data/access_15.rda")), compress = "xz")
```

@@

```{r}
#by DA and mean:
access_15_DA <- access_15 |>
  left_join(HAM_census_21 |> st_drop_geometry() |> select(c("GeoUID", "Population", "Dwellings", "LICO_AT_TOTAL", "COMMUNITY_")), 
            by=c("GeoUID")) |> 
  group_by(GeoUID) |>
  summarise(copp_iscDA_med = median(copp_isc), #the median number of opportunities (all) that can be reached from each parcel in a DA
            copp_iscDA_sd = sd(copp_isc), #the SD
            V_iscDA = sum(V_isc), #the number of total spatially available opportunities (all) that can be reached from each parcel in a DA
            vv_iscDA = sum(V_isc)/first(Population), #the per-capita number of total spatially available opportunities (all) that can be reached from each parcel in a DA
            Population = first(Population),
            Dwellings = first(Dwellings),
            LICO_AT_TOTAL = first(LICO_AT_TOTAL),
            COMMUNITY_ = first(COMMUNITY_)) |>
  tidyr::drop_na(Population)

@@

```{r}
#calculating entropy, here I do it by the 4 categories for each parcel and then aggregate to DA. Another way is by the 14 specific categories. For now - keep it simple and just focus on the 4 categories. This assumes that all specific types are the SAME within categories.
step1 <- access_15  |> group_by(from_id) |> 
  summarise(OPPi = sum(copp_isc, na.rm=T),
            Vi = sum(V_isc))

step2 <- access_15 |> left_join(step1, by="from_id")

step2 <- step2 |> mutate(
  top1_copp = (copp_isc) / OPPi * log((copp_isc) / OPPi),
  top1_copp = ifelse(is.nan(top1_copp), 0, top1_copp),

  top1_V= (V_isc) / Vi * log((V_isc) / Vi),
  top1_V = ifelse(is.nan(top1_V), 0, top1_V),

  entropy_copp_isc = sum(top1_copp) / -log(4),
  entropy_V_isc = sum(top1_V) / -log(4))  |> 
  group_by(GeoUID) |>
  summarise(entropy_copp_isc = mean(entropy_copp_isc, na.rm=T) + sd(entropy_copp_isc, na.rm=T),
            entropy_V_isc = mean(entropy_V_isc, na.rm=T) + sd(entropy_V_isc, na.rm=T))

@@

```{r}
#the number of opportunities from each categories
all_cats <- access_15_bycategory_DA |> 
  mutate(copp_all = sum(copp_ic_med, na.rm=T) + sum(copp_ic_sd, na.rm=T)) |>
  group_by(GeoUID) |>
         summarise(summed_copp_all = sum(copp_all),
                   summed_Vic = sum(V_ic, na.rm=T),
                   Population = mean(Population),
                   Care_Access_Status1 = case_when(summed_copp_all < 4 ~ "[0-04 des.) Not caring",
                                 summed_copp_all >= 4 & summed_copp_all < 25 ~ "[04-25 des.) Caring", 
                                 summed_copp_all >= 25 ~ "[25+ des.) Very caring"),
                   Care_Access_Status2 = case_when(all(copp_all >= 2 & copp_all < 4) ~ "- complete (2 to 4 options per cat.)", 
                                                   all(copp_all >= 4) ~ "- very complete (4 or more options per cat.)"),
                   entropy_copp_isc = first(entropy_copp_isc),
                   entropy_V_isc = first(entropy_V_isc)) |>
  mutate(Care_Access_Status2 = ifelse(is.na(Care_Access_Status2), "", Care_Access_Status2),
         Care_Access_Status = paste(Care_Access_Status1,Care_Access_Status2, sep=" "),
         vvi = summed_Vic/Population)

@@

```{r message=FALSE}
#tmap_mode("plot")
all_cats_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(all_cats) +
  tm_polygons("Care_Access_Status",
              border.alpha = 0.2,
              palette = c("darkred", "orangered", "orange", "yellow", "forestgreen", "lightgreen"),
              title = "Number of care destinations spatially accessible within 15 min. walk",
              legend.is.portrait = TRUE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("right", "top"), size=1.0)+
  tm_layout(bg.color = "grey",
            panel.label.bg.color = "white",
            panel.label.color = "black",
            legend.outside = TRUE,
            legend.position = c("center", "bottom"),
            legend.outside.position = "bottom",
            legend.outside.size = 0.2,
            legend.width = 12
            # legend.title.size = 0.7,
            # legend.text.size = 0.6
            )

tmap_save(all_cats_copp_plot, file= file.path(here("figures/all_cats_copp_plot.tiff")), dpi=600)
```

```{r}
all_cats_entropy_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(all_cats) +
  tm_polygons("entropy_copp_isc",
              border.alpha = 0.2,
              palette = "Reds",
              title = "Diversity of care destinations spatially accessible within 15 min. walk",
              legend.is.portrait = TRUE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("right", "top"), size=1.0)+
  tm_layout(bg.color = "grey",
            panel.label.bg.color = "white",
            panel.label.color = "black",
            legend.outside = TRUE,
            legend.position = c("center", "bottom"),
            legend.outside.position = "bottom",
            legend.outside.size = 0.2,
            legend.width = 12)

tmap_save(all_cats_entropy_copp_plot, file= file.path(here("figures/all_cats_entropy_copp_plot.tiff")), dpi=600)
```

```{r}
all_cats_vvi_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(all_cats) +
  tm_polygons("vvi",
              border.alpha = 0.2,
              palette = c("orangered", "forestgreen"),
              breaks = c(0, 0.003804, 0.022174),
              labels = c("Below Median", "Median and above"),
              title = "Per capita spatially available care destinations within 15 min. walk",
              legend.is.portrait = TRUE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("right", "top"), size=1.0)+
  tm_layout(bg.color = "grey",
            panel.label.bg.color = "white",
            panel.label.color = "black",
            legend.outside = TRUE,
            legend.position = c("center", "bottom"),
            legend.outside.position = "bottom",
            legend.outside.size = 0.2,
            legend.width = 12
            )

tmap_save(all_cats_vvi_plot, file= file.path(here("figures/all_cats_vvi_plot.tiff")), dpi=600)
```

@@

```{r}
#adding entropy labels
all_cats <- all_cats |> 
  mutate(bivariate_entropy = 
           case_when( ((Care_Access_Status == "[04-25 des.) Caring - complete (2 to 4 options per cat.)"  |
                       Care_Access_Status =="[04-25 des.) Caring - very complete (4 or more options per cat.)" |
                         Care_Access_Status =="[25+ des.) Very caring - very complete (4 or more options per cat.)") &
                         entropy_copp_isc >= 0.7418) ~ "Caring, complete and amenity diverse",  #the median diversity,

                     ((Care_Access_Status == "[04-25 des.) Caring - complete (2 to 4 options per cat.)"  |
                       Care_Access_Status =="[04-25 des.) Caring - very complete (4 or more options per cat.)" |
                         Care_Access_Status =="[25+ des.) Very caring - very complete (4 or more options per cat.)") &
                        entropy_copp_isc < 0.7418) ~ "Caring, complete but not amenity diverse",

                     ((Care_Access_Status == "[0-04 des.) Not caring " |
                       Care_Access_Status == "[04-25 des.) Caring ")  &
                        entropy_copp_isc >= 0.7418) ~ "Not caring or not complete, but amenity diverse",

                     ((Care_Access_Status == "[0-04 des.) Not caring "  |
                       Care_Access_Status == "[04-25 des.) Caring ")  &
                        entropy_copp_isc < 0.7418) ~ "Not caring, complete, or amenity diverse"))

@@

```{r}
all_cats_bivariate_entropy_copp_plot <- tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(all_cats) +
  tm_polygons("bivariate_entropy",
              border.alpha = 0.2,
              palette = c("yellowgreen","yellow","blue", "#FFFFDF"),
              title = "Diversity of care destinations spatially accessible within 15 min. walk",
              legend.is.portrait = TRUE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("right", "top"), size=1.0)+
  tm_layout(bg.color = "grey",
            panel.label.bg.color = "white",
            panel.label.color = "black",
            legend.outside = TRUE,
            legend.position = c("center", "bottom"),
            legend.outside.position = "bottom",
            legend.outside.size = 0.2,
            legend.width = 12)

tmap_save(all_cats_bivariate_entropy_copp_plot, 
          file= file.path(here("figures/all_cats_bivariate_entropy_copp_plot.tiff")),
          dpi=600)
```

```{r}
care_categories_plot <- 
 tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(Community_Boundaries, bbox=Community_Boundaries) + tm_polygons(col = "white", border.col = "black", lwd=1)+
  tm_shape(care_dest) + tm_dots("Care_Category_Specific", palette = RColorBrewer::brewer.pal(n = 4, name = 'Dark2'), size=0.025, title = "")+
  tm_facets("Care_Category", nrow = 2, free.scales= TRUE)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 20)) +
  tm_compass(position = c("left", "top"), size=0.9) +
  tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 0.5, legend.bg.color = "white", legend.bg.alpha = 0.4,
            panel.labels = c("1,265 Dependent-centric destinations", 
                             "158 Errand-centric destinations", 
                             "381 Grocery-centric destinations", 
                             "421 Health-centric destinations"),
            panel.label.size = 0.8)

tmap_save(care_categories_plot, 
          file= file.path(here("figures/care_categories_plot.tiff")),
          dpi=600)
```
```{r}
parcels_DA_plot <- 
 tm_shape(ham_bay, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0) +
  tm_shape(hydro_p_LakeOntario, bbox=Community_Boundaries) + tm_polygons(col="skyblue", border.alpha = 0)+
  tm_shape(access_15_DA) +
  tm_polygons("parcel_count",
              border.alpha = 0.2,
              palette = "Reds",style = "quantile", n=4,
              title = "Count of residential parcels",
              legend.is.portrait = TRUE) +
  tm_shape(Community_Boundaries) + tm_polygons(alpha=0)+
  tm_scale_bar(position = c("left", "bottom"), breaks = c(0, 1, 5, 10, 15)) +
  tm_compass(position = c("left", "top"), size=1.0)+
   tm_layout(bg.color = "grey", panel.label.bg.color = "white", legend.position = c("right", "top"), legend.text.size = 0.5,legend.title.size = 0.7,  legend.bg.color = "white", legend.bg.alpha = 0.4)

@@

pop_descriptive_plots <- tmap_arrange(parcels_DA_plot, poprate_DA_plot, LICOAT_DA_plot, nrow=3)

tmap_save(pop_descriptive_plots, 
          file= file.path(here("figures/pop_descriptive_plots.tiff")) , asp = 1/0.6,
          dpi=600)
```
